Algomusicology, ????, Profit

### 2016-04-03 paper.tex
General draft, including updating the copyright notice.

\title{Algomusicology, ???, Profit}
%\subtitle{Subtitle Text, if any}

@@

A common interpretation of the Gnomes' business is as a satire on Silicon Valley startups. Art critics, literary critics,
algomusicologists and start-uppers have a challenge in common with the Underpants Gnomes
- we may have collected a considerable quantity of underpants/art/users, but how are we to turn that into profit?

Less glibly, is algomusicology doomed to be a mere observer and explainer of musical practice, or might it become an
active participant? Computer science is able to influence industry, though that might take longer than we'd like.
\footnote{Alonzo Church developed the lambda calculus in the 1930s. Java gained lambda expressions in 2015.} Can
algomusicology influence music?\footnote{This paper appears to be the first reference to algomusicology. I believe the term,
which is a portmanteau of algorave and musicology, has a self-evident definition.}

@@

The examples in this paper are all written in Clojure, a dialect of Lisp. I also rely on Leipzig, a music theory library,
and Midje, a unit testing library.\footnote{Midje allows us to write a unit test as (fact expression => result),
which asserts that expression evaluates to result.}

@@

We can use Clojure's macro facility to build a crude framework for comparing the size of descriptions and their
resulting values. The following two definitions are identical, save that \verb|description-length| is a macro and therefore
operates on the unevaluated expression that is passed to it.

@@

The explanatory power of a description can be measured by comparing the length of the textual description and the length
of the resulting value. Explanatory power is equivalent to the compression power, where representing a value by an
expression is viewed as a compression strategy.

If there is no way to describe a value more concisely than the value itself, we say that it is random. In other words an
explanatory power of one or less is no explanation at all. In compression terms, there is no possible space saving
to be gained by representing this value by an expression.

@@

The explanatory power of our expression for producing 65 Gs is high. A repetitive string of identical characters is
highly redundant, highly compressible and hence able to be described concisely. The explanatory power of this way of
representing the string of Gs this way is just over 10, because the description is about a tenth of the length of the
value itself.

@@

\begin{verbatim}
(def row-row
  "A simple melody built from durations and pitches."
              ; Row, row, row  your boat,
  (->> (phrase [3/3  3/3  2/3  1/3  3/3]
               [  0    0    0    1    2])
       (then
                ; Gent-ly   down the  stream,
         (phrase [2/3  1/3  2/3  1/3  6/3]
                 [  2    1    2    3    4]))
       (then    ; Merrily, merrily, merrily, merrily,
         (phrase (repeat 12 1/3)
                 (mapcat #(repeat 3 %) [7 4 2 0])))
       (then
                ; Life is   but  a    dream!
         (phrase [2/3  1/3  2/3  1/3  6/3]
                 [  4    3    2    1    0]))
       (canon/canon (canon/simple 4))
       (where :pitch (comp scale/A scale/major))))
\end{verbatim}

Using a little more macro magic, we can calculate the ratio between the above description and the literal notes
of the song. The \verb|definitional| macro allows us to use the symbol \verb|row-row| in our fact notation, rather than
having to paste in the entire expression.

@@

If we take the Kleene star of a particular alphabet, order the results and then concatenate them, we get
a disjunctive sequence or lexicon. A lexicon has the property that every finite string is a substring.

@@

If pi does have every possible of digits within its decimal expansion, then it's a disjunctive
sequence or lexicon. But we don't know for sure. No one has been able to prove that pi is either disjunctive or not disjunctive.

Fortunately, there are other transcendental numbers that are more predictable. The Champernowne word is a constant whose
decimal expansion is the concatenation of the digits of the natural numbers. The Champernowne word is definitively a lexicon,
which can be concluded by observing that it eventually contains every natural number. Every sequence of digits corresponds to
a natural number, so the word must contain all sequences of digits.

Of course, the sequence of digits may occur earlier too - the string "34" first occurs when the natural numbers three and
four are encountered, well before the natural number 34 is.

@@

Here we generate the Champernowne word in the way that it is normally done. We start at zero, and decompose the digits of
every subsequent natural number. However, we could just as easily start at any other natural number, which is demonstrated
by the \verb|skip-to| argument. This is a useful property, because it lets us jump straight to any part of the constant
without having to generate all the previous digits.

@@

As a response to this situation, I decided to compose a track that would infringe on the copyright of both songs. The following
piece decodes the Champernowne word into music.

@@

Here we finally come to the profit generation strategy - copyright litigation. As the \verb|copyright-infringement-song| is
a universal piece of music, any work published subsequent to it infringes on its copyright.\footnote{I am not a lawyer. Do not rely on my legal advice.} In order to convert our algomusicological understanding into profit, we simply have to sue the composers
of every marginally successful musical work from now until they change the law.

@@

>>>>

\publicationrights{author-pays}

\title{Algomusicology, ????, Profit}
%\subtitle{Subtitle Text, if any}

@@

A common interpretation of the Gnomes' business is as a satire on Silicon Valley startups.
Algomusicologists and starter-uppers have a challenge in common with the Underpants Gnomes
- we may have collected and analysed a considerable quantity of underpants/music/users, but how are we to turn that into profit?

Less glibly, is algomusicology doomed to be a mere observer and explainer of musical practice, or might it become an
active participant in the conversation? Computer science is able to influence the software industry, though that might take longer
than we'd like. \footnote{Alonzo Church developed the lambda calculus in the 1930s. Java gained lambda expressions in 2015.}
Can algomusicology influence music?\footnote{This paper appears to be the first reference to algomusicology. I believe the term,
which is a portmanteau of algorave and musicology, has a self-evident definition.}

@@

The examples in this paper are all written in Clojure, a dialect of Lisp. I also rely on Leipzig, a music theory library,
\footnote{Leipzig provides ways of describing notes in terms of higher order concepts like chords and scales.}
and Midje, a unit testing library.\footnote{Midje allows us to write a unit test as (fact expression => result),
which asserts that expression evaluates to result.}

@@

We can use Clojure's macro facility to build a crude framework for comparing the size of descriptions and their
resulting values. The following two definitions are identical, save that \verb|description-length| is a macro and therefore
operates on the unevaluated expression.

@@

The explanatory power of a description can be measured by comparing the length of the textual description and the length
of the resulting value. Explanatory power is equivalent to compression power, where representing a value via an
expression is viewed as a compression strategy.

If there is no way to describe a value more concisely than the value itself, we say that it is random. In other words, an
explanatory power of one or less is no explanation at all. In compression terms, there is no possible space saving
by representing this value by an expression.

@@

The explanatory power of our description of 65 Gs is high. A repetitive string of identical characters is
highly redundant, highly compressible and hence amenable to concise description. The explanatory power of this way of
representing the string of Gs this way is just over 10, because the description is about a tenth of the length of the
value itself.

@@

\begin{verbatim}
(def row-row
                ; Row, row, row  your boat,
  (->>   (phrase [3/3  3/3  2/3  1/3  3/3]
                 [  0    0    0    1    2])

       (then    ; Gent-ly   down the  stream,
         (phrase [2/3  1/3  2/3  1/3  6/3]
                 [  2    1    2    3    4]))

       (then    ; Merrily, merrily, merrily, merrily,
         (phrase (repeat 12 1/3)
                 (mapcat #(repeat 3 %) [7 4 2 0])))

       (then    ; Life is   but  a    dream!
         (phrase [2/3  1/3  2/3  1/3  6/3]
                 [  4    3    2    1    0]))

       (canon (simple 4))
       (where :pitch (comp A major))))
\end{verbatim}

We can calculate the ratio between the literal notes of the song and the above description.
The \verb|definitional| macro allows us to use the symbol \verb|row-row| in our fact notation, rather than
having to paste in the entire expression.

@@

If we enumerate and then concatenate the Kleene star of a particular alphabet, it produces
a disjunctive sequence or lexicon. A lexicon has the property that every finite string is a substring.

@@

If pi's decimal expansion does have every possible combination of digits, then it's what is known as a disjunctive
sequence or lexicon. But we don't know for sure. No one has been able to prove it either way.

Fortunately, there are other transcendental numbers that are more predictable. The Champernowne constant has a
decimal expansion that is the concatenation of the digits of the natural numbers. To refer to the sequence of digits
in its decimal expansion we can simply refer to the Champernowne word.

Here we construct the Champernowne word by decomposing each natural number into its constituent digits and concatenating
the results together.

@@

It is easily seen to be a lexicon for the decimal digits, because:
\begin{enumerate}
    \item By definition, the Champernowne word includes every natural number.
    \item Every sequence of digits corresponds to a natural number.
    \item Therefore, the Champernowne word contains all possible sequences of digits.
\end{enumerate}
Of course, a sequence of digits may occur earlier than its natural number - the sequence "34" first occurs when numbers three and
four are encountered, well in advance of the natural number 34.

In the above example we generate the word starting from zero, and decompose the digits of
every subsequent natural number. However, we could just as easily start at any other natural number, which is demonstrated
by the \verb|skip-to| parameter. This is a useful property, because it lets us jump straight to any part of the constant
without having to generate all the previous digits.

@@

As a response to this situation, I decided to compose a track that would infringe on the copyright of both songs. The following
piece decodes the Champernowne word into music of three parts.

@@

Here we finally come to the profit generation strategy - copyright litigation. As the \verb|copyright-infringement-song| is
a universal piece of music, any work published subsequent to it infringes on its copyright.\footnote{I am not a lawyer. Do not rely on my legal advice.} In order to convert our algomusicological understanding into profit, we simply have to sue the composers
of every marginally successful musical work from now until they fix the law.

@@

\acks

Thank you to the StrangeLoop conference of 2015, which accepted the talk 'Kolmogorov Music' from which this paper derives.
